---
title: "RNA-seq Analysis of UV-Induced RASSF9 Expression in Human Skin"
author: "Samson Kosemani"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: journal
    highlight: tango
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6,
  fig.align = "center"
)

# Your analysis code continues here...

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6,
  fig.align = "center"
)

# Your analysis code continues here...

Abstract
Analysis of GEO dataset GSE245747 to investigate UVB-induced expression of RASSF9 and related genes in human skin explants using RNA-seq data.

Introduction
This study examines the transcriptional response to UVB radiation in human skin, focusing on:
  
  RASSF9 expression dynamics

RASSF family member co-regulation

Potential involvement in UV stress response pathways

Methods
Data Acquisition

# Install necessary packages
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
if (!require("GEOquery", quietly = TRUE))
  BiocManager::install("GEOquery")
if (!require("tidyverse", quietly = TRUE))
  install.packages("tidyverse")
if (!require("ggplot2", quietly = TRUE))
  install.packages("ggplot2")
if (!require("ggpubr", quietly = TRUE))
  install.packages("ggpubr")
if (!require("DESeq2", quietly = TRUE))
  BiocManager::install("DESeq2")
if (!require("airway", quietly = TRUE))
  BiocManager::install("airway")

library(GEOquery)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(DESeq2)

Dataset Description
GEO Accession: GSE245747
Title: UVB induces RASSF9 expression in human skin
Technology: RNA-seq
Samples: Human skin explants, UVB-treated vs control, multiple time points

# Create a personal library folder if it doesn't exist
if (!dir.exists(Sys.getenv("R_LIBS_USER"))) {
  dir.create(Sys.getenv("R_LIBS_USER"), recursive = TRUE)
}

# Add it to your library paths
.libPaths(c(Sys.getenv("R_LIBS_USER"), .libPaths()))

# Now install packages to your user directory
install.packages(c("tidyverse", "ggplot2", "ggpubr"), lib = Sys.getenv("R_LIBS_USER"))

# For Bioconductor packages
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager", lib = Sys.getenv("R_LIBS_USER"))

BiocManager::install(c("GEOquery", "DESeq2"), lib = Sys.getenv("R_LIBS_USER"))
